---
title: "General Social Survey"
author: "Aubrey Garcia"
date: "Summer 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  

First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: Gun ownership and cappun are unrelated
$H_A$: Gun ownership and cappun are related

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

CAPPUN is a catgorical varible
OWNAGUN is a catgorical varible,
We want CAPPUN~OWNGUN, 
so  use the Cat~Cat helper file.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results



<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata, type="percent")

```

The graph shows that among those who own a gun ABOUT  70% of them are in favor of CAPPUN and from among those who don't own a gun and are in favor of CAPUUN only 45% are in favor.

#### Numerical Descriptive Results

```{r}
table1 <- xtabs(~OWNGUN + CAPPUN , data=GSSdata)
rowPerc(table1)
colPerc(table1)
```
Numerical summary shows that 75% of gun owners favor CAPPUN while about 25% oppose CAPPUN.


The table shows that 

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->



According to the graph, it shows that 

### Inferential Results

  The P-value for the chi-square test was p=0.0202, so this means we reject the null hypothesis and conclude that cappun and and gun ownership are related. 
  
  The P-vaule for for the fisher exact test was p=0.01651and the odds ration was 2:671:1.
  
  So the odds ratio 2:671:1 means that people who don't own a gun are 2.271 times likely to oppose capitial punishment.

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



 The P-value for the chi-square test was p=0.0202, hence we reject the null hypothesis and conclude that gun owner ship and capital punishment are not independent.
 
 
  The P-vaule for for the fisher exact test was p=0.01651 hence we reject the null hypothesis, as with the chi-square test, and furthermore the odds ratio says that gunowns are 2.3 times more likely to favor capital punishment than those who don't own a gun.
 

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is not a difference of education between those who did or did not vote in 2012.

$H_A$: There is a difference of education betwen those who did and did not vote in 2012.

## Methods
Education is categorical varible
Votes is numerical variable
So we use the Numerical~Cat helper file.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results


```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

The Graph shows that those who voted in 2012 tend to have higher education that those who didnt vote.

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```


According to the table there is a slight difference in education of those who did and didn't vote. Those who did vote had about 16 years of education which is slightly higer than those with 13 years of education  of 13.4 which are the ones that didn't vote.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```


According to the T-test the p-vaule was 0.0000001248, which is less than .05 so we reject the null hypothesis. Signifying there isa difference on those who did and did't vote based on their education levels.

```{r}
fisher.test(table1)
```
The P value in the chi-square test was 0.0202 which is less than .05 so we reject the null hypothesis

In the fishers exact test the p-vaule was 0.01651 which is less than 0.5 so we reject the null hypothesis the test also shows an odds ratio of 2.271.
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).


$H_0$:A persons performance on the wordsum test is not related to their level of schooling.

$H_A$: A persons performance on the wordsum test is related to their level of schooling

## Methods

Performance on the wordsum test is a numerical varible.

Level of schooling is a catergorical varible.

So we use the Numerical~Cat helper file.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
plot(WORDSUM~DEGREE,data=GSSdata)
```
According to the bar plot, it shows that there was a difference on the outcome of someones WORDSUM performance, with their level of education being a factor in the result.
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

According to the table there was a slight difference in the means in the peoples performance on the WORDSUM test based on their level of education.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
model1 <- aov(WORDSUM~DEGREE,data = GSSdata)
summary.aov(model1)
```

According to the table the P-value was 0.0000002, which is less than .05 so therefore the null hypothesis.

l
l
'l;
l
l;
;
;
;


The TukeyHSD test shows there is no difference in highschool and less than highschool.There is a differnce between jr high and high school. There is a difference between bachlors and high school.There is a differnce in graduate and high school. There is no differnce between high school and gratuate. There is no differnce between jr college and graduate, There is no differnce betweem bachlors and graduate. 


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

 In question two, after being asked weather there was a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)? After much testing, the p-value was 0.0000001248, which is less than .5 so we rejected the null hypothesis that stated There was not a difference of education between those who did or did not vote in 2012. Concluding there was a difference of those who voted based on their education. On question 3 The issue was to determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE). After testing the p-value was 0.00000025, which is less than .5 so therefore the null hypothesis that stated a persons performance on the wordsum test is not related to their level of schooling was rejected. Concluding that a person performance on the wordsum is related to their level of schooling.


<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

### Question 1: Is gun ownership an indepedent of opinion on capital punishment. The summary resuults show that they are not independent. Thefore Gun owners are 2.3 more times likely to be in favor of gun ownership, that those who do not.


### Question 2: 
