---
title: "General Social Survey"
author: "Aubrey Garcia"
date: "Summer 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: Gun ownership and cappun are unrelated
$H_A$: Gun ownership and cappun are related

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

CAPPUN is a catgorical varible
OWNAGUN is a catgorical varible,
so i use the Cat~Cat helper file.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata, type="percent")

```

The graph shows that

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
table1 <- xtabs(~CAPPUN + OWNGUN , data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

The table shows that..

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->



According to the graph, it shows that 

### Inferential Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

  The P-vaule for the chi-squre test was p=0.0202, so this means we reject the null hypothesis and conclude that ...
  
  The P-vaule for for the fisher exact test was p=..... and the odds ration was yyyy..
  
  So the odds ratio 2:671:2:

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is not a difference of education between those who did or did not vote in 2012.

$H_A$: There is a difference of education btween those who did and did not vote in 2012.

## Methods
Education is categorical vaible
Votes is numerical varible
So we use the Num~Cat helper file.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results


```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

The Graph shows there is a difference in education and to those....

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```


According to the table there is a slight difference in education of those who did and didn't vote. Those who did vote had a mean of 15.72 which is slightly higer than the mean of 13.4 which are the ones that didn't vote.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```
According to the T-test the p-vaule was 0.0000001248, which is less than .5 so we reject the null hypothesis...


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).


$H_0$:A persons performance on the wordsum test is not related to their level of schooling.

$H_A$: A persons performance on the wordsum test is related to their level of schooling

## Methods

Performance on the wordsum test is a numerical varible.

Level of schooling is a catergorical varible.

So we use the Numerical~Cat helper file.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
plot(WORDSUM~DEGREE,data=GSSdata)
```
According to the bar plot...
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

According to the table...

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
model1 <- aov(WORDSUM~DEGREE,data = GSSdata)
summary.aov(model1)
```

According to the table...

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

As shown above in question one, after various testing results there is a significant relationship between gun ownership and opinion on the death. The results supported that statement, giving us a p- vaule of p=0.0202. Thus rejecting the Null Hypothesis which stated that Gun ownership and cappun are unrelated. In question two, after being asked weather there was a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)? After much testing, the p-value was 0.0000001248, which is less than .5 so we rejected the null hypothesis that stated There was not a difference of education between those who did or did not vote in 2012. Concluding there was a difference of those who voted based on their education. On question 3 The issue was to determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE). After testing the p-value was 0.00000025, which is less than .5 so therefore the null hypothesis that stated a persons performance on the wordsum test is not related to their level of schooling was rejected. Concluding that a person performance on the wordsum is related to their level of schooling.


<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

